### 2014-12-13 paper.tex

  When people think of measuring gerrymandering, the first kind of metric that comes to mind is generally one that tries to measure the compactness of a district.  In fact, there are a number of ways of measuring compactness; some are better than others, but many measure different types of compactness.  There are three major types: those that measure the dispersion of the district from its center, those that measure some form of boundary complexity, and those that are based on population and population density.
% TODO: maybe this wants a figure to show some of the metrics visually? I would make one but my laptop crashes when I try to use qgis. At least convex hull probably wants this, maybe also circumscribing/inscribed circle although those matter less since we never computed them.

  The very simplest metrics for compactness simply measure the length and width of the district as compared to that of a circle or rectangle of the same area.  There are several variations on this method, but they are all fairly crude.  Other metrics compare the area of the district to the area of the smallest circumscribing circle, or the largest inscribed circle, or to the convex hull of the district -- the smallest convex shape containing the entire district.  Niemi~et~al.~\cite{niemi} list a number of variations on these metrics.  An alternate method of computing dispersion is by computing the moment of inertia of the area of the district.  The concept is borrowed from physics, in which rotating bodies have a moment of inertia that defines how their mass is distributed: a bicycle wheel has a higher moment of inertia than a ball of the same mass because its mass is further from its axis, and thus the wheel is more inclined to keep rotating once started.  The moment of area simply assumes that the district has equal ``mass'' everywhere, and is defined as
  \[I = \iint \left((x-x_0)^2 + (y-y_0)^2\right) dx\, dy \]
  where $(x_0, y_0)$ is the center of mass of the district (weighted by area).  More dispersed districts have higher moments of inertia, because the average square distance of a point in the district to the center is larger.  In order to ensure that the metric does not depend on the size of the state, and to make it range from 0 to 1 with 0 being the least compact, we in fact use the measure $A/\sqrt{2\pi I}$.  This metric seems to have been first proposed by Kaiser.~\cite{kaiser}

>>>>

  When people think of measuring gerrymandering, the first kind of metric that comes to mind is generally one that tries to measure the compactness of a district.  In fact, there are a number of ways of measuring compactness; some are better than others, but many measure different types of compactness.  There are three major types: those that measure the dispersion of the district from its center, those that measure some form of boundary complexity, and those that are based on population and population density.

  \begin{figure}
    \begin{center}
      \parbox{\textwidth}{\centering \includegraphics[width=0.5\textwidth]{convexhull.png} \\ Convex Hull}
      \parbox{\textwidth}{\centering \parbox{0.9\textwidth}{\caption{Convex hull of a shape. The original shape is in blue, and the convex hull encompasses both that area and the yellow sections.\label{f:convexhull}}}}
    \end{center}
  \end{figure}

  The very simplest metrics for compactness simply measure the length and width of the district as compared to that of a circle or rectangle of the same area.  There are several variations on this method, but they are all fairly crude.  Other metrics compare the area of the district to the area of the smallest circumscribing circle, or the largest inscribed circle, or to the convex hull of the district -- the smallest convex shape containing the entire district, as shown in Figure~\ref{f:convexhull}.  Niemi~et~al.~\cite{niemi} list a number of variations on these metrics.  An alternate method of computing dispersion is by computing the moment of inertia of the area of the district.  The concept is borrowed from physics, in which rotating bodies have a moment of inertia that defines how their mass is distributed: a bicycle wheel has a higher moment of inertia than a ball of the same mass because its mass is further from its axis, and thus the wheel is more inclined to keep rotating once started.  The moment of area simply assumes that the district has equal ``mass'' everywhere, and is defined as
  \[I = \iint \left((x-x_0)^2 + (y-y_0)^2\right) dx\, dy \]
  where $(x_0, y_0)$ is the center of mass of the district (weighted by area).  More dispersed districts have higher moments of inertia, because the average square distance of a point in the district to the center is larger.  In order to ensure that the metric does not depend on the size of the state, and to make it range from 0 to 1 with 0 being the least compact, we in fact use the measure $A/\sqrt{2\pi I}$.  This metric seems to have been first proposed by Kaiser.~\cite{kaiser}
